dispRity manual

### 2016-05-29 doc/dispRity-manual.Rmd
Added code to the Rmd docs (removed redundant Rnw)

\

\
\
\
\
\

    ## [1] 50 48

    ##                    [,1]          [,2]        [,3]       [,4]      [,5]
    ## Cimolestes   -0.5319679  0.1117759259  0.09865194 -0.1933148 0.2035833
    ## Maelestes    -0.4087147  0.0139690317  0.26268300  0.2297096 0.1310953
    ## Batodon      -0.6923194  0.3308625215 -0.10175223 -0.1899656 0.1003108
    ## Bulaklestes  -0.6802291 -0.0134872777  0.11018009 -0.4103588 0.4326298
    ## Daulestes    -0.7386111  0.0009001369  0.12006449 -0.4978191 0.4741342
    ## Uchkudukodon -0.5105254 -0.2420633915  0.44170317 -0.1172972 0.3602273

\
\

    ## [1] 99 97

` `

    ##                  [,1]       [,2]       [,3]       [,4]        [,5]
    ## Cimolestes -0.6082437 -0.0323683 0.08458885 -0.4338448 -0.30536875
    ## Maelestes  -0.5730206 -0.2840361 0.01308847 -0.1258848  0.06123611

` `

    ##            [,1]      [,2]       [,3]       [,4]       [,5]
    ## n48 -0.05529018 0.4799330 0.04118477 0.04944912 -0.3558830
    ## n49 -0.13067785 0.4478168 0.11956268 0.13800340 -0.3222785

\
\

    ##             FAD  LAD
    ## Adapis     37.2 36.8
    ## Asioryctes 83.6 72.1
    ## Leptictis  33.9 33.3
    ## Miacis     49.0 46.7
    ## Mimotona   61.6 59.2
    ## Notharctus 50.2 47.0

\
\
`         `

@@

\
`     `\
`           `\
`     `\
\
\
`   `\
\
\
`     `\
\
\

    ##   series  n observed  mean  2.5%   25%   75% 97.5%
    ## 1    120  5    1.494 1.200 0.756 0.854 1.484 1.550
    ## 2    105 11    1.725 1.646 1.481 1.614 1.686 1.770
    ## 3     90 18    1.820 1.775 1.681 1.747 1.806 1.849
    ## 4     75 19    1.923 1.814 1.681 1.778 1.865 1.901
    ## 5     60 20    1.946 1.859 1.751 1.833 1.894 1.925
    ## 6     45 14    1.930 1.849 1.742 1.827 1.883 1.951
    ## 7     30 10    1.910 1.796 1.570 1.774 1.866 1.935
    ## 8     15 10    1.910 1.801 1.594 1.782 1.853 1.914
    ## 9      0 10    1.910 1.815 1.588 1.788 1.865 1.929

\
\
\
`     `

    ##              Df Sum Sq Mean Sq F value Pr(>F)    
    ## series        8  34.47   4.309   279.5 <2e-16 ***
    ## Residuals   891  13.74   0.015                   
    ## ---
    ## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

@@

\

@@

\

    ## [1] "dispRity"

\

    ## [1] "data"      "disparity" "elements"  "series"    "call"

\

    ## Disparity measurements across 9 series for 99 elements 
    ## Series:
    ## 120, 105, 90, 75, 60, 45 ...
    ## Disparity calculated as: c(median, centroids) for 97 dimensions.
    ## Data was split using continuous method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

Note however, that it is always possible to recall the full object using
the argument `all=TRUE`:

\
`   `\

@@

\
`     `\

    ## Disparity measurements across 2 series for 24 elements 
    ## Series:
    ## 120, 75.
    ## Disparity calculated as: c(median, centroids) for 97 dimensions.
    ## Data was split using continuous method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

\

    ##      120      105       90       75       60       45       30       15        0 
    ## 1.493570 1.725280 1.820292 1.922531 1.945952 1.930306 1.910203 1.910203 1.910203

\
`   `\
`      `

    ## List of 1
    ##  $ 120: num [1:100] 1.52 0.786 1.555 1.288 0.891 ...

@@

\
`  `

    ##   series  n observed
    ## 1    120  5   -0.005
    ## 2    105 11   -0.010
    ## 3     90 18   -0.010
    ## 4     75 19   -0.001
    ## 5     60 20    0.005
    ## 6     45 14    0.013
    ## 7     30 10    0.017
    ## 8     15 10    0.017
    ## 9      0 10    0.017

\
`   `\
\
\
`  `

    ##   series  n observed
    ## 1    120  5   -0.005
    ## 2    105 11   -0.010
    ## 3     90 18   -0.010
    ## 4     75 19   -0.001
    ## 5     60 20    0.005
    ## 6     45 14    0.013
    ## 7     30 10    0.017
    ## 8     15 10    0.017
    ## 9      0 10    0.017

@@

\
\
`     `\
`    `\
\
`  `\
`         `\
`          `\
\
\
\
`  `\
\
\

    ## [1] "dispRity"

    ## 2 custom series for 50 elements 
    ## Series:
    ## Group.crown, Group.stem.

\

    ## [1] "data"     "elements" "series"

\

    ## List of 2
    ##  $ Group.crown: num [1:30, 1:48] 0.3079 0.6531 0.5089 -0.1652 -0.0419 ...
    ##   ..- attr(*, "dimnames")=List of 2
    ##   .. ..$ : chr [1:30] "Dasypodidae" "Bradypus" "Myrmecophagidae" "Todralestes" ...
    ##   .. ..$ : NULL
    ##  $ Group.stem : num [1:20, 1:48] -0.739 -0.68 -0.511 -0.477 -0.473 ...
    ##   ..- attr(*, "dimnames")=List of 2
    ##   .. ..$ : chr [1:20] "Daulestes" "Bulaklestes" "Uchkudukodon" "Kennalestes" ...
    ##   .. ..$ : NULL

\

    ##  chr [1:50] "Cimolestes" "Maelestes" "Batodon" "Bulaklestes" "Daulestes" ...

\
\

    ## [1] "custom"      "Group.crown" "Group.stem"

@@

\
`  `\
`         `\
`        `\
`           `\
\
\

    ## 3 custom series for 50 elements 
    ## Series:
    ## Group.crown, Group.stem, Class.Mammalia.

@@

\
`      `\
`         `

    ## 3 discrete series for 50 elements 
    ## Series:
    ## 120-80, 80-40, 40-0.

In this example, the taxa where split inside each timebin according to
their age. Note however that some taxa might span between some time bins
and should be included in more than one. This is possible by providing a
table containing the first and last occurrence data:

\
\
`     `\
`              `\

    ## 3 discrete series for 50 elements 
    ## Series:
    ## 120-80, 80-40, 40-0.

@@

\
`      `\
`      `

    ## 3 discrete series for 50 elements 
    ## Series:
    ## 133.51104-89.00736, 89.00736-44.50368, 44.50368-0.

@@

\
\
`     `\
`               `\
`     `\

    ## 4 continuous series for 99 elements 
    ## Series:
    ## 120, 80, 40, 0.

\
\
`      `\
`         `

    ## 4 continuous series for 99 elements 
    ## Series:
    ## 133.51104, 89.00736, 44.50368, 0.

@@

\
` `

    ## Bootstrapped ordinated matrix with 50 elements 
    ## 1
    ## Data was bootstrapped 1000 times, using the full bootstrap method.

\
\

@@

\
`   `

    ## Bootstrapped ordinated matrix with 50 elements 
    ## 1
    ## Data was bootstrapped 1000 times, using the single bootstrap method.

\
`      `

    ## Bootstrapped ordinated matrix with 50 elements 
    ## 1
    ## Data was bootstrapped 20 times, using the full bootstrap method.
    ## Data was fully rarefied (down to 3 elements).

\
`      `

    ## Bootstrapped ordinated matrix with 50 elements 
    ## 1
    ## Data was bootstrapped 20 times, using the full bootstrap method.
    ## Data was rarefied with a maximum of 6, 7, 8 and 3 elements

\
`   `

    ## Bootstrapped ordinated matrix with 50 elements 
    ## 1
    ## Data was bootstrapped 1000 times, using the full bootstrap method. 
    ## The 6 last axis were removed from the original ordinated data.

<span>`## Or with a fix value (50%)`</span>\
`   `

    ## Bootstrapped ordinated matrix with 50 elements 
    ## 1
    ## Data was bootstrapped 1000 times, using the full bootstrap method. 
    ## The 35 last axis were removed from the original ordinated data.

@@

\
`        `\
\
\
`     `\
`     `\
\
\

    ## Bootstrapped ordinated matrix with 50 elements 
    ## Series:
    ## Group.crown, Group.stem.
    ## Data was split using custom method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.
    ## Data was fully rarefied (down to 3 elements).

    ## Bootstrapped ordinated matrix with 50 elements 
    ## Series:
    ## 120-80, 80-40, 40-0.
    ## Data was split using discrete method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

    ## Bootstrapped ordinated matrix with 99 elements 
    ## Series:
    ## 120, 80, 40, 0.
    ## Data was split using continuous method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.

@@

`  `

    ## Disparity measurements across 1 series for 50 elements 
    ## 1
    ## Disparity calculated as: mean for 48 dimensions.

\
\
\
\
`   `

    ## Disparity measurements across 1 series for 50 elements 
    ## 1
    ## Disparity calculated as: c(sum, variances) for 48 dimensions.

\
\
\
`   `

    ## Disparity measurements across 1 series for 50 elements 
    ## 1
    ## Disparity calculated as: c(prod, centroids) for 48 dimensions.

\
\
\
\
`   `\
`   `

    ## Disparity measurements across 1 series for 50 elements 
    ## 1
    ## Disparity calculated as: c(total.range, centroids) for 48 dimensions.

\
\
\
\
`     `\
\
\
\
\
\
`     `\
`  `

    ## [1] TRUE

@@

\
`     `\
\
\
`     `\
`     `

@@

    ##   series  n observed  mean  2.5%   25%   75% 97.5%
    ## 1 120-80  8    1.203 1.103 0.895 1.057 1.178 1.236
    ## 2  80-40 27    1.344 1.322 1.275 1.306 1.339 1.366
    ## 3   40-0 16    1.353 1.323 1.265 1.304 1.346 1.368

\
\
\
\
\
`        `

    ##   series  n observed  sd  6% 94%
    ## 1 120-80  8      1.2 0.1 1.0 1.2
    ## 2  80-40 27      1.3 0.0 1.3 1.4
    ## 3   40-0 16      1.4 0.0 1.3 1.4

\
\
`   `

    ## Disparity calculated as: c(median, centroids) for 48 dimensions.
    ## Data was split using discrete method.
    ## Data was bootstrapped 100 times, using the full bootstrap method.
    ##   series  n observed  mean  2.5%   25%   75% 97.5%
    ## 1 120-80  8    1.203 1.103 0.895 1.057 1.178 1.236
    ## 2  80-40 27    1.344 1.322 1.275 1.306 1.339 1.366
    ## 3   40-0 16    1.353 1.323 1.265 1.304 1.346 1.368

\
\

    ##        series  n observed  mean  2.5%   25%   75% 97.5%
    ## 41 Group.stem 15       NA 1.229 1.146 1.210 1.253 1.282
    ## 42 Group.stem 16       NA 1.232 1.154 1.217 1.250 1.279
    ## 43 Group.stem 17       NA 1.233 1.174 1.218 1.251 1.281
    ## 44 Group.stem 18       NA 1.236 1.175 1.219 1.255 1.282
    ## 45 Group.stem 19       NA 1.239 1.173 1.226 1.256 1.286
    ## 46 Group.stem 20     1.27 1.240 1.185 1.225 1.258 1.280

@@

\
`          `\
\
\
`   `\
\
\
\
`      `\
\
`         `

@@

\
`  `\
\
`      `

@@

\
`  `\
\
\
`      `

@@

\
`  `\
\
\
`   `\

@@

\
`  `\
\
`     `\
`           `\
`           `\
`         `\
`         `

@@

\
`  `\
\
\
`   `\
\
\
`  `\
`      `

@@

\
\
`  `

    ##                                 t       df      p.value mean of x mean of y
    ## Group.crown - Group.stem 42.25372 164.8723 2.303646e-90   1.36798  1.239652

\
`     `

    ## $`Group.crown - Group.stem`
    ## 
    ##  Welch Two Sample t-test
    ## 
    ## data:  data[[list_of_comp[[1]]]] and data[[list_of_comp[[2]]]]
    ## t = 42.254, df = 164.87, p-value < 2.2e-16
    ## alternative hypothesis: true difference in means is not equal to 0
    ## 95 percent confidence interval:
    ##  0.1223319 0.1343251
    ## sample estimates:
    ## mean of x mean of y 
    ##  1.367980  1.239652

\
\
`  `\
`         `

    ##             W      p.value
    ## 120 - 80    0 7.457311e-34
    ## 80 - 40  4448 5.334298e-01
    ## 40 - 0   1600 2.961960e-16

\
`     `

    ## Call:
    ##    test(formula = data ~ series, data = data)
    ## 
    ## Terms:
    ##                   series Residuals
    ## Sum of Squares  33.50243  10.79609
    ## Deg. of Freedom        3       396
    ## 
    ## Residual standard error: 0.1651147
    ## Estimated effects may be unbalanced

\
`       `\
` `

    ## [1] "aov" "lm"
    ##              Df Sum Sq Mean Sq F value Pr(>F)    
    ## series        3   33.5  11.167   409.6 <2e-16 ***
    ## Residuals   396   10.8   0.027                   
    ## ---
    ## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

\
\
\
\
`    `\
`         `

    ##                bhatt.coeff
    ## 120-80 - 80-40  0.04472136
    ## 120-80 - 40-0   0.03741657
    ## 80-40 - 120-80  0.04472136
    ## 40-0 - 120-80   0.03741657

@@

\

    ## [1] 139.074

\
` `

    ##    ages     elements
    ## 1 90.00    Daulestes
    ## 2 90.00  Bulaklestes
    ## 3 90.00 Uchkudukodon
    ## 4 77.85  Kennalestes
    ## 5 77.85   Asioryctes

\
`    `

    ##    ages     elements
    ## 1 0.647    Daulestes
    ## 2 0.647  Bulaklestes
    ## 3 0.647 Uchkudukodon
    ## 4 0.560  Kennalestes
    ## 5 0.560   Asioryctes

\
\
`       `

    ##        ages     elements
    ## 1 0.3528637    Daulestes
    ## 2 0.3528637  Bulaklestes
    ## 3 0.3528637 Uchkudukodon
    ## 4 0.4402271  Kennalestes
    ## 5 0.4402271   Asioryctes

>>>>

```{r}
## Loading the package
library(dispRity)

## Setting the random seed for repeatability
set.seed(123)

## Loading the ordinated matrix containing 50 taxa
data(BeckLee_mat50)
dim(BeckLee_mat50)
head(BeckLee_mat50[,1:5])

## Loading another ordinated matrix containing 50 tips + 49 nodes
data(BeckLee_mat99)
dim(BeckLee_mat99)
head(BeckLee_mat99[,1:5], 2)
tail(BeckLee_mat99[,1:5], 2)

## Loading a list of first and last occurrence data
data(BeckLee_ages)
head(BeckLee_ages)

## Loading the phylogeny
data(BeckLee_tree)
plot(BeckLee_tree, cex = 0.8) ; nodelabels(cex = 0.8) ; axisPhylo(root = 140)
```

@@

```{r}
## Splitting the data
sliced_data <- time.series(BeckLee_mat99, BeckLee_tree, method = "continuous",
    model = "acctran", time = rev(seq(from=0, to=130, by=15)),
    FADLAD = BeckLee_ages)

## Bootstrapping the data
bootstrapped_data <- boot.matrix(sliced_data, 100)

## Calculating disparity
median_centroids <- dispRity(bootstrapped_data, metric = c(median, centroids))

## Summarising the results
summary(median_centroids)
plot(median_centroids)

## Testing the effect of time on disparity
summary(test.dispRity(median_centroids, test = aov, comparisons = "all"))
```

@@

```{r}
str(median_centroids)
## That's a more than 4000 lines of output!
```

@@

```{r}
## What is the class of the median_centroids object?
class(median_centroids)

## What is in the object?
names(median_centroids)

## Summarising it using the S3 method print.dispRity:
median_centroids
## This displays some basic information about the object content
@

\noindent Note however, that it is always possible to recall the full object using the argument \texttt{all=TRUE}:

<<dispRity class3, eval=FALSE>>=
## Displaying the full object
print(median_centroids, all = TRUE)
## But that's 155091 lines!
```

@@

```{r}
## Extracting some specific series from the disparity object
series_1_and_4 <- get.dispRity(median_centroids, what = c(1,4))
series_1_and_4

## The observed disparity
extract.dispRity(median_centroids)

## The distribution of the bootstrapped disparity scores for the first series
str(extract.dispRity(get.dispRity(median_centroids, what = 1),
    observed = FALSE))
```

@@

```{r}
## Disparity measured as the mean value of the ordinated matrix
summary(dispRity(sliced_data, metric = mean))

## A function for measuring the sum divided by the length (the mean!)
sum.divide.by.length <- function(X) sum(X)/length(X)

## Disparity measured as the mean divided by the length
summary(dispRity(sliced_data, metric = sum.divide.by.length))
## Giving the exact same results!
```

@@

```{r}
## We want to separate the species around the node 71 (see phylogeny above).
## All the descendant of these node are crown and all the ancestors are stem.
crown <- extract.clade(BeckLee_tree, node = 71)$tip.label
stem <- drop.tip(BeckLee_tree, tip = crown)$tip.label
## We then have to feed this information in a data frame with one column
factors <- as.data.frame(
    matrix(data = c(rep("crown", length(crown)), rep("stem", length(stem))),
    ncol = 1, dimnames = list(c(crown, stem), "Group")))

## We then can use the customised series function to create the two series
## i.e. the two regions of the ordinated space
crown_stem <- cust.series(BeckLee_mat50, factors)

## This created a dispRity object containing two series: crown and stem
class(crown_stem)
crown_stem

## This object contains three elements
names(crown_stem)
## With "data" being the list sub-matrices
str(crown_stem$data)
## "elements" being the list of taxa in the original ordinated matrix
str(crown_stem$elements)
## and finally, "series" containing information on the series type (custom)
## and names (crown and stem)
crown_stem$series
```

@@

```{r}
## Adding all mammals to the a new series (along with crown and stem as before)
factors2 <- as.data.frame(
    matrix(data = c(rep("crown", length(crown)), rep("stem", length(stem)),
    rep("Mammalia", Ntip(BeckLee_tree))), ncol = 2,
    dimnames = list(c(crown, stem), c("Group", "Class")), byrow = FALSE))

## Creating the three series: Group.stem/crown and Class.Mammalia
cust.series(BeckLee_mat50, factors2)
```

@@

```{r}
## Generating three time bins containing the data present every 40 Ma
time.series(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
    time = c(120, 80, 40, 0))

```
In this example, the taxa where split inside each timebin according to
their age. Note however that some taxa might span between some time bins
and should be included in more than one. This is possible by providing a
table containing the first and last occurrence data:

```{r}
## Generating three time bins containing the data present every 40 Ma, including
## taxa that might span between time bins.
time_bins <- time.series(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = c(120, 80, 40, 0), FADLAD = BeckLee_ages)
time_bins

```

@@

```{r}
## Automatically generate three equal length bins:
time.series(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
    time = 3)
```

@@

```{r}
## Generating four time slices every 40 million years assuming a gradual
## evolution model
time_slices <- time.series(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "gradual", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)
time_slices
## Note that in the same way as for the discrete method, we can also
## automatically generate the slices
time.series(data = BeckLee_mat99, tree = BeckLee_tree, method = "continuous",
    model = "gradual", time = 4)
```

@@

```{r}
## Bootstrapping a matrix
boot.matrix(data = BeckLee_mat50)
## As we can see, the output is also a dispRity object that is summarised
## automatically, and gives information on the data as well as the number of
## bootstraps and the bootstraps methods.
```

@@

```{r}
## Bootstrapping with the single bootstrap method
boot.matrix(BeckLee_mat50, boot.type = "single")

## Bootstrapping with the full rarefaction
boot.matrix(BeckLee_mat50, bootstraps = 20, rarefaction = TRUE)

## Or with a set number of rarefaction levels
boot.matrix(BeckLee_mat50, bootstraps = 20, rarefaction = c(6:8,3))

## And removing the last axis (default)
boot.matrix(BeckLee_mat50, rm.last.axis = TRUE)

## Or with a fix value (50%)
boot.matrix(BeckLee_mat50, rm.last.axis = 0.5)
```

@@

```{r}
## Bootstrap and full rarefaction on the crown/stem series
crown_stemBS <- boot.matrix(crown_stem, bootstraps = 100, rarefaction = TRUE)

## Bootstrap on the time binning/slicing series
time_binsBS <- boot.matrix(time_bins, bootstraps = 100)
time_slicesBS <- boot.matrix(time_slices, bootstraps = 100)

## Note that all these objects are of class dispRity
crown_stemBS
time_binsBS
time_slicesBS
```

@@

```{r}
dispRity(BeckLee_mat50, metric = mean)
## This defines disparity as the mean value of the morphospace
## (using base::mean).

## It is also possible to combine multiple functions:
dispRity(BeckLee_mat50, metric = c(sum, variances))
## Defining disparity as the sum (base::sum) of the variances
## (dispRity::variances) of each dimension of the morphospace.

dispRity(BeckLee_mat50, metric = c(prod, centroids))
## For the product (base::prod) of the distances between each elements and the
## centroid of the morphospace (dispRIty::centroids).

## Or user defined ones:
total.range <- function(X) abs(range(X)[1]-range(X)[2])
dispRity(BeckLee_mat50, metric = c(total.range, centroids))
## For the range (total.range, user defined) of the distances between each
## elements and the centroid of the morphospace (dispRity::centroids).

## Or more complex ones:
vars1 <- dispRity(BeckLee_mat50, metric = c(sd, variances, var))
## For the standard deviation (stats::sd) of the variance (dispRity::variances)
## of each column of the variance/covariance matrix (stats::var).

## Note that the order of the function is not important since the levels of
## each function are automatically detected by the dispRity algorithm.
vars2 <- dispRity(BeckLee_mat50, metric = c(variances, var, sd))
all(summary(vars1) == summary(vars2))
```

@@

```{r}
## Disparity in crown and stem mammals:
disp_crown_stemBS <- dispRity(crown_stemBS, metric = c(median, centroids))

## Disparity through time:
disp_time_binsBS <- dispRity(time_binsBS, metric = c(median, centroids))
disp_time_slicesBS <- dispRity(time_slicesBS, metric = c(median, centroids))

```

@@

```{r}
summary(disp_time_binsBS)
## By default, the 50 and 95 quantiles and the mean are calculated.
## Note that the function also displays the observed disparity
## (non-bootstrapped).

## These arguments can be changed easily as follow
summary(disp_time_binsBS, quantile = 88, cent.tend = sd, rounding = 1)

## If on happens to forgot what was calculated in the summarised object,
## it is possible to recall the different steps details using recall:
summary(disp_time_binsBS, recall = TRUE)

## Note that the information with each different number of elements is
## displayed for rarefied data:
tail(summary(disp_crown_stemBS))
## The observed disparity is NA for rarefied data.
```

@@

```{r}
## Graphical options
quartz(width = 15, height = 5) ; par(mfrow = (c(1,3)), bty = "n")

## Plotting the disparity for both groups
plot(disp_crown_stemBS, type = "box")

## Same plot but with the number of elements on the right and using the
## "polygon" plot type.
plot(disp_crown_stemBS, type = "polygon", elements = TRUE)
## Rarefied version of the same plot (20 elements per series)
plot(disp_crown_stemBS, type = "polygon", elements = TRUE, rarefaction = 20)
```

@@

```{r}
## Graphical options
par(bty = "n")
## The rarefaction curves
plot(disp_crown_stemBS, elements = TRUE, rarefaction = "plot")

```

@@

```{r}
## Graphical options
par(bty = "n")
## The disparity-through-time data in a time-binned way (with lines rather than
## boxes) with the observed data as crosses
plot(disp_time_binsBS, type = "lines", observed = TRUE)
```

@@

```{r}
## Graphical options
par(bty = "n")

## First, plotting the continuous data with a y axis limit
plot(disp_time_slicesBS, type = "continuous")
## Note that the type = "continuous" argument is not mandatory here.
```

@@

```{r}
## Graphical options
par(bty = "n")
## A plot with many options!
plot(disp_time_slicesBS, quantile = c(seq(from=10, to=100, by=10)),
    cent.tend = sd, type = "continuous", elements = "log",
    col = c("black", rainbow(10)), ylab = c("Disparity", "log(Diversity)"),
    xlab = "Time (in in units from past to present)", time.series = FALSE,
    observed = TRUE, main = "Many options...")
```

@@

```{r}
## Graphical options
par(bty = "n")

## First, plotting the continuous data with a y axis limit
plot(disp_time_slicesBS, ylim = c(0.5,2))

## Second, adding the discrete data
par(new = TRUE)
plot(disp_time_binsBS, discrete_type = "line", ylim=c(0.5,2), xlab="", ylab="")
```

@@

```{r}
## Performing a pairwise t-test to the test a difference in disparity between
## crown and stem mammals (note that "pairwise" is default)
test.dispRity(disp_crown_stemBS, test = t.test)

## The same test but with the detailed output
test.dispRity(disp_crown_stemBS, test = t.test, details = TRUE)

## A wilcoxon rank test applied to sliced disparity the data sequentially with
## the Bonferroni correction.
test.dispRity(disp_time_slicesBS, test = wilcox.test,
    comparisons = "sequential", correction = "bonferroni")

## An analysis of variance (aov) applied to the same data
test.dispRity(disp_time_slicesBS, test = aov, comparisons = "all")

## Note that in this case, the output is the regular aov output:
slice_aov <- test.dispRity(disp_time_slicesBS, test = aov, comparisons = "all")
class(slice_aov) ; summary(slice_aov)

## Measuring the overlap between distributions in the time bins (using the
## implemented Bhattacharyya Coefficient function - see ?bhatt.coeff) with a
## personalised series comparisons and some additional arguments to pass to the
## test function (bw = bw.nrd)
test.dispRity(disp_time_binsBS, test = bhatt.coeff, bw = bw.nrd,
    comparisons = list(c(1,2), c(1,3), c(2,1), c(3,1)))

```

@@

```{r}
## This tree has a root age
BeckLee_tree$root.time
## So we can get the age of each tips and nodes directly
head(tree.age(BeckLee_tree), 5)
## But we can also decide to make the age relative (between 1 and 0)
head(tree.age(BeckLee_tree, age = 1), 5)
## Or even relative, but from the root (i.e. how far are the nodes/tips
## from the root)
head(tree.age(BeckLee_tree, age = 1, order = "present"), 5)
```
